---
title: "Problem Set 1_Applied Stats_Dongli Wu_21362988"
output: html_notebook
---


```{r}
# Question 1

# Make the hypotheses for the test:

## Null hypothesis(Ho): the empirical distribution of the observed data matches the specified PDF.
## Alternative hypothesis(H1): at least one value doesn't match the specified distribution.


# set seed for random number generation

set.seed(123)

# generate 1000 random cauchy density varialbes 
y <-rcauchy(1000, location = 0, scale = 1)
y

# write the function to find the largest absolute difference between the empirical distirbution (use "ecdf(y)(y)") and the theoretical normal distribution (use "pnorm(y)") across all x position.

D <- y[which.max(abs(ecdf(y)(y)-pnorm(y,mean(y),sd = sd(y))))]

# set up the observed data as data frame.
df <-data.frame(x = y)

# plot the Empirical distribution, Theoretical distribution and also mark the maximum difference.

library(ggplot2)

ggplot(df, aes(x)) +
  stat_ecdf(aes(colour = "Empirical")) +
  stat_function(fun = pnorm, args = list(mean = mean(y), sd = sd(y)),
                aes(colour = "Theoretical")) +
  annotate(
    "segment",
    x = D, xend = D,
    y = ecdf(y)(D), yend = pnorm(D, mean(y), sd(y)),
    size =1
  )

# The maxinum difference shows on the gragh, D = 0.3 (=0.4-0.1)

# Now we can use "ks.test()" to get P-value and test if the D value the same result as above

ks.test(y,"pnorm",mean(y),sd(y))

## D = 0.30395, D-value equals to the result above.
## p-value < 2.2e-16, P-value more less than 0.05, so the null hypothesis could be rejected. The empirical distribution of the observed data doesn't match the theoretical distribution.


```

```{r}
#Question 2

# set up the data

set.seed(123)
data <-data.frame(x = runif(200, 1, 100))
data$y <- 0 + 2.75*data$x +rnorm(200, 0, 1.5)

# write the function of the OLS regression with θ=(β, σ)
linear.lik <- function(theta, y, x){
  n <- nrow(x)
  k <- ncol(x)
  beta <- theta[1:k]
  sigma2 <- theta[k+1]^2
  e <- y-x%*%beta
  logl <- -0.5*n*log(2*pi)-0.5*n*log(sigma2)-((t(e)%*%e)/(2*sigma2))
  return(-logl)
}



# it creates a surface at different values of β and σ

surface <- list()
k <- 0
for (beta in seq (0, 5, 0.1)){
   for (sigma in seq(0.1, 5, 0.1)){
     k <- k + 1
     logl <- linear.lik(theta = c(0, beta, sigma), y = data$y, x = cbind(1, data$x))
     surface[[k]] <- data.frame(beta = beta, sigma = sigma, logl = -logl)
   }
}
surface <- do.call(rbind, surface)
library(lattice)

wireframe(logl ~ beta*sigma, surface, shade = TRUE)

# use "optim()" to find the parameters of the maximum point on the surface

linear.MLE <- optim(fn=linear.lik, par = c(1,1,1), hessian = TRUE, y = data$y, x = cbind(1, data$x), method = "BFGS")

linear.MLE$par

##[1]  0.118865  2.747869 -1.441841
## the result explores the parameters,  β=2.75 and σ^2=118

summary(lm(y ~x, data))

##Coefficients:
##            Estimate Std. Error t value            Pr(>|t|)
##(Intercept) 0.118004   0.218720    0.54                0.59
##x           2.747882   0.003781  726.81 <0.0000000000000002
               
##(Intercept)    
##x           ***
##---

# Conclusion: lm turns out the equivalent results to the Newton-Raphson algotiyhm.
```

```{r}

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

